---
title: "Case-Shiller 3-Tier City Index Assessment"
author: "Bingjun Li"
date: "3/23/2018"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# the warnings caused by the high difference in ARIMA don't affect the resulting models.
knitr::opts_chunk$set(warning = FALSE)

# load necessary library
library(forecast) # for prediction
library(dygraphs) # for visulization

# load all csv file into dataframe
temp = list.files(pattern="*.csv")
for (i in 1:length(temp)) {
  assign(strtrim(temp[i], (nchar(temp[i]) - 4)), read.csv(temp[i]))
}

# create a function to trim to the time series object that is a multiplier of 12 monthes
yearTrim <- function(tsobj){
  # to obtain the number of years and the number of left-over months
  numy <- floor(length(tsobj) / 12)
  extra <- length(tsobj) - numy * 12
  
  # to get the start year and month of the object
  start <- start(tsobj)
  
  # to trim the time series object
  result <- window(tsobj, start = c(start[1], start[2] + extra))
  
  #return the trimmed time series object
  return(result)
}

# the cross-validation implied here is a rolling basis
# for example, the first training group is year 1 and test group is year 2
# then, the second training group is year 1 and 2 and test group is year 3
# the training group keeps rolling up each iteration and test group is always the next 12 months

# define a function to find a optimal model for input data
# tsobj input is time series object
nextYearPred <- function(tsobj){
  
  # to get the start time
  stime <- start(tsobj)
  
  # to obtain the number of trainning groups
  numy = length(tsobj) / 12 - 1
  
  # to create one numeric vector for storing mae value and one list for ARIMA parameters
  models <- vector("list", numy)
  maes <- vector("numeric", numy)
  
  # loop through every year
  for (i in 2:numy){
  
    # fit the model on the data
    tsobjfc <- auto.arima(window(tsobj, stime, c(stime[1], stime[2] + i * 12 - 1)), D = 1)

    # store model parameters
    models[[i]] <- tsobjfc$arma
    
    # generate the prediction for the test group
    pred <- forecast(tsobjfc, 12)
    
    # store the mean absolute errors (MAEs)
    maes[i] <- mean(abs(pred$mean - window(tsobj, c(stime[1] + i, stime[2]), c(stime[1] + i, stime[2] + 11))))
  }
  
  # using mae criterion for the best fitting model
  para <- models[[which.min(maes[2:numy]) + 1]]
  
  # fit the optimal model to the whole data set
  # and predict for the next whole year
  tsobjfc.a <- arima(tsobj, order = c(para[1], para[6], para[2]), seasonal = list(order = c(para[3], para[7], para[4]), period = para[5]), method = "ML")
  prediction <- forecast(tsobjfc.a, (24 - extra))
  # return the forecast object
  return(prediction)
}

# convert Boston Home Index data (3 tiers) into time series object
boxrhtnsats <- ts(BOXRHTNSA[,2], start = c(1987, 1), frequency = 12)
boxrltnsats <- ts(BOXRLTNSA[,2], start = c(1987, 1), frequency = 12)
boxrmtnsats <- ts(BOXRMTNSA[,2], start = c(1987, 1), frequency = 12)

# trim the time series objects to be a multiplier of 12 monthes
boxrhtnsats <- yearTrim(boxrhtnsats)
boxrltnsats <- yearTrim(boxrltnsats)
boxrmtnsats <- yearTrim(boxrmtnsats)

# run the function for high tier in Boston
pred.bosht <- nextYearPred(boxrhtnsats)$mean

# repeat the modelling process on other two tiers

# for low tier
pred.boslt <- nextYearPred(boxrltnsats)$mean

# for middle tier
pred.bosmt <- nextYearPred(boxrmtnsats)$mean

# combine the original data and the prediction into one data frame 
boxrnsa <- cbind(boxrhtnsats, boxrmtnsats, boxrltnsats, pred.bosht, pred.bosmt, pred.boslt)

# use dygraphs package to plot the data and the prediction for three tiers in one graph
dygraph(boxrnsa, main = "Home Price Index for Boston, Massachusetts") %>%
  dyAxis("x", label = "Time", drawGrid = FALSE, pixelsPerLabel = 40, axisLabelFormatter = "function(d){return d.getFullYear();}") %>%
  dyAxis("y", label = "Price Index") %>%
  dySeries("boxrhtnsats", label = "High Tier", strokeWidth = 3) %>%
  dySeries("boxrmtnsats", label = "Middle Tier", strokeWidth = 3) %>%
  dySeries("boxrltnsats", label = "Low Tier", strokeWidth = 3) %>%
  dySeries("pred.bosht", label = "Prediction for High Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dySeries("pred.bosmt", label = "Prediction for Middle Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dySeries("pred.boslt", label = "Prediction for Low Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dyOptions(colors = RColorBrewer::brewer.pal(3, "Dark2")) %>%
  dyShading(from = "2018-1-1", to = "2018-12-1") %>%
  dyLegend(width = 600)

# to compute the difference betwen the index at two different time
# a compareInd function was created. Input is a TS object, 
# the prior time, as a year and month integer vector, such as c(2018, 1) 
# and the time gap between in month
compareInd <- function(tsobj, time1, timegap){
  tmp <- window(tsobj, time1)
  res <- (tmp[timegap + 1] - tmp[1])/ tmp[1]
  return(res)
}

# compare the index for all three tiers between the most current and one years ago
compareInd(boxrhtnsats, c(2016,12), 12)
compareInd(boxrmtnsats, c(2016,12), 12)
compareInd(boxrltnsats, c(2016,12), 12)

# compare the predicted index in 2018.12 and the index in 2017.12
(pred.bosht[12] - boxrhtnsats[372])/ boxrhtnsats[372]
(pred.bosmt[12] - boxrmtnsats[372])/ boxrmtnsats[372]
(pred.boslt[12] - boxrltnsats[372])/ boxrltnsats[372]
```
In each metro area, residential homes were divided into 3 tiers, high tier, middle tier and low tier. From the plot for Boston, MA, it can be observed that the home price indexes for all 3 tiers have exceeded the highest position before 2008 financial crisis by the end of 2017. The high, middle and low tiers of residential homes' year over year growth rates by December 2017 are 4.25%, 6.73%, and 7.8%. Using seasonal ARIMA model, we can obtain the prediction for 2018. By the end of 2018, the predicted year over year growth rates for 3 tiers are 3% for the high tier, 5.61% for the middle tier and 7.21% for the low tier. It can be clearly observed that the low tier is the backbone of the Boston market growth in 2018. Both the low tier and the middle tier is predicted to maintain roughly the same growth rate as that in 2017, but the growth of the high tier is predicted to slow down in 2018.

```{r}
knitr::opts_chunk$set(warning = FALSE)

# convert Chicago Home Index data (3 tiers) into time series object
chxrhtnsats <- ts(CHXRHTNSA[,2], start = c(1992, 1), frequency = 12)
chxrltnsats <- ts(CHXRLTNSA[,2], start = c(1992, 1), frequency = 12)
chxrmtnsats <- ts(CHXRMTNSA[,2], start = c(1992, 1), frequency = 12)

# trim the time series objects to be a multiplier of 12 monthes
chxrhtnsats <- yearTrim(chxrhtnsats)
chxrltnsats <- yearTrim(chxrltnsats)
chxrmtnsats <- yearTrim(chxrmtnsats)

# predict for high tier
pred.chiht <- nextYearPred(chxrhtnsats)$mean

# predict for low tier
pred.chilt <- nextYearPred(chxrltnsats)$mean

# predict for middle tier
pred.chimt <- nextYearPred(chxrmtnsats)$mean

# combine the original data and the prediction into one data frame 
chxrnsa <- cbind(chxrhtnsats, chxrmtnsats, chxrltnsats, pred.chiht, pred.chimt, pred.chilt)

# use dygraphs package to plot the data and the prediction for three tiers in one graph
dygraph(chxrnsa, main = "Home Price Index for Chicago, Illinois") %>%
  dyAxis("x", label = "Time", drawGrid = FALSE, pixelsPerLabel = 40, axisLabelFormatter = "function(d){return d.getFullYear();}") %>%
  dyAxis("y", label = "Price Index") %>%
  dySeries("chxrhtnsats", label = "High Tier", strokeWidth = 3) %>%
  dySeries("chxrmtnsats", label = "Middle Tier", strokeWidth = 3) %>%
  dySeries("chxrltnsats", label = "Low Tier", strokeWidth = 3) %>%
  dySeries("pred.chiht", label = "Prediction for High Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dySeries("pred.chimt", label = "Prediction for Middle Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dySeries("pred.chilt", label = "Prediction for Low Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dyOptions(colors = RColorBrewer::brewer.pal(3, "Dark2")) %>%
  dyShading(from = "2018-1-1", to = "2018-12-1") %>%
  dyLegend(width = 600)

# compare the index for all three tiers between the most current and one years ago
compareInd(chxrhtnsats, c(2016,12), 12)
compareInd(chxrmtnsats, c(2016,12), 12)
compareInd(chxrltnsats, c(2016,12), 12)

# compare the predicted index in 2018.12 and the index in 2017.12
(pred.chiht[12] - chxrhtnsats[312])/ chxrhtnsats[312]
(pred.chimt[12] - chxrmtnsats[312])/ chxrmtnsats[312]
(pred.chilt[12] - chxrltnsats[312])/ chxrltnsats[312]
```
It can be observed that the home price indexes for all 3 tiers in Chicago, IL have experienced significant growth, though it hasn't exceeded the highest position before 2008 financial crisis by the end of 2017. The high, middle and low tiers of residential homes' year over year growth rates by December 2017 are 0.38%, 4.92%, and 9.63%. It's clear that the low tier residential home is the main growth factor of Chicago market in 2017. Using the seasonal ARIMA model, the predicted year over year growth rates for 3 tiers by the end of 2018 are 0.38% for the high tier, 4.64% for the middle tier and 6.38% for the low tier. It can be seen that the growth rate for all 3 tiers homes will fall in 2018 but still maintain the growing tendency.

```{r}
# convert Los Angeles Home Index data (3 tiers) into time series object
laxrhtnsats <- ts(LXXRHTNSA[,2], start = c(1987, 1), frequency = 12)
laxrltnsats <- ts(LXXRLTNSA[,2], start = c(1987, 1), frequency = 12)
laxrmtnsats <- ts(LXXRMTNSA[,2], start = c(1987, 1), frequency = 12)

# trim the time series objects to be a multiplier of 12 monthes
laxrhtnsats <- yearTrim(laxrhtnsats)
laxrltnsats <- yearTrim(laxrltnsats)
laxrmtnsats <- yearTrim(laxrmtnsats)

# predict for high tier
pred.laht <- nextYearPred(laxrhtnsats)$mean

# predict for low tier
pred.lalt <- nextYearPred(laxrltnsats)$mean

# predict for middle tier
pred.lamt <- nextYearPred(laxrmtnsats)$mean

# combine the original data and the prediction into one data frame 
laxrnsa <- cbind(laxrhtnsats, laxrmtnsats, laxrltnsats, pred.laht, pred.lamt, pred.lalt)

# use dygraphs package to plot the data and the prediction for three tiers in one graph
dygraph(laxrnsa, main = "Home Price Index for Los Angeles, California") %>%
  dyAxis("x", label = "Time", drawGrid = FALSE, pixelsPerLabel = 40, axisLabelFormatter = "function(d){return d.getFullYear();}") %>%
  dyAxis("y", label = "Price Index") %>%
  dySeries("laxrhtnsats", label = "High Tier", strokeWidth = 3) %>%
  dySeries("laxrmtnsats", label = "Middle Tier", strokeWidth = 3) %>%
  dySeries("laxrltnsats", label = "Low Tier", strokeWidth = 3) %>%
  dySeries("pred.laht", label = "Prediction For High Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dySeries("pred.lamt", label = "Prediction For Middle Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dySeries("pred.lalt", label = "Prediction For Low Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dyOptions(strokeWidth = 3, colors = RColorBrewer::brewer.pal(3, "Dark2")) %>%
  dyShading(from = "2018-1-1", to = "2018-12-1") %>%
  dyLegend(width = 600)

# compare the index for all three tiers between the most current and one years ago
compareInd(laxrhtnsats, c(2016,12), 12)
compareInd(laxrmtnsats, c(2016,12), 12)
compareInd(laxrltnsats, c(2016,12), 12)

# compare the predicted index in 2018.12 and the index in 2017.12
(pred.laht[12] - laxrhtnsats[372])/ laxrhtnsats[372]
(pred.lamt[12] - laxrmtnsats[372])/ laxrmtnsats[372]
(pred.lalt[12] - laxrltnsats[372])/ laxrltnsats[372]
```
It can be observed that the home price indexes for all 3 tiers in Los Angeles, CA have experienced significant growth and has been around the highest position before 2008 financial crisis by the end of 2017. The high, middle and low tiers of residential homes' year over year growth rates by December 2017 are 6.53%, 7.35%, and 10.7%, which is one the highest across the country. Using the seasonal ARIMA model, the predicted year over year growth rates for 3 tiers by the end of 2018 are 7.61% for the high tier, 7.8% for the middle tier and 16.16% for the low tier, all of which are expected to exceed to the highest mark achieved before 2008 financial crisis. It can be seen that the growth rate for all 3 tiers homes in 2018 will exceed that in 2017 and the future of the Los Angeles home market seems very bright.

```{r}
knitr::opts_chunk$set(warning = FALSE)

# convert New York Home Index data (3 tiers) into time series object
nyxrhtnsats <- ts(NYXRHTNSA[,2], start = c(1987, 1), frequency = 12)
nyxrltnsats <- ts(NYXRLTNSA[,2], start = c(1987, 1), frequency = 12)
nyxrmtnsats <- ts(NYXRMTNSA[,2], start = c(1987, 1), frequency = 12)

# trim the time series objects to be a multiplier of 12 monthes
nyxrhtnsats <- yearTrim(nyxrhtnsats)
nyxrltnsats <- yearTrim(nyxrltnsats)
nyxrmtnsats <- yearTrim(nyxrmtnsats)

# predict for high tier
pred.nyht <- nextYearPred(nyxrhtnsats)$mean

# predict for low tier
pred.nylt <- nextYearPred(nyxrltnsats)$mean

# predict for middle tier
pred.nymt <- nextYearPred(nyxrmtnsats)$mean

# combine the original data and the prediction into one data frame 
nyxrnsa <- cbind(nyxrhtnsats, nyxrmtnsats, nyxrltnsats, pred.nyht, pred.nymt, pred.nylt)

# use dygraphs package to plot the data and the prediction for three tiers in one graph
dygraph(nyxrnsa, main = "Home Price Index for New York, New York") %>%
  dyAxis("x", label = "年份", drawGrid = FALSE, pixelsPerLabel = 40, axisLabelFormatter = "function(d){return d.getFullYear();}") %>%
  dyAxis("y", label = "Price Index") %>%
  dySeries("nyxrhtnsats", label = "High Tier", strokeWidth = 3) %>%
  dySeries("nyxrmtnsats", label = "Middle Tier", strokeWidth = 3) %>%
  dySeries("nyxrltnsats", label = "Low Tier", strokeWidth = 3) %>%
  dySeries("pred.nyht", label = "Prediction for High Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dySeries("pred.nymt", label = "Prediction for Middle Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dySeries("pred.nylt", label = "Prediction for Low Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dyOptions(strokeWidth = 3, colors = RColorBrewer::brewer.pal(3, "Dark2")) %>%
  dyShading(from = "2018-1-1", to = "2018-12-1") %>%
  dyLegend(width = 600)

# compare the index for all three tiers between the most current and one years ago
compareInd(nyxrhtnsats, c(2016,12), 12)
compareInd(nyxrmtnsats, c(2016,12), 12)
compareInd(nyxrltnsats, c(2016,12), 12)

# compare the predicted index in 2018.12 and the index in 2017.12
(pred.nyht[12] - nyxrhtnsats[372])/ nyxrhtnsats[372]
(pred.nymt[12] - nyxrmtnsats[372])/ nyxrmtnsats[372]
(pred.nylt[12] - nyxrltnsats[372])/ nyxrltnsats[372]
```
It can be observed that the home price indexes for all 3 tiers in New York, NY have experienced steady growth, but hasn't been back to the high mark before 2008 financial crisis by the end of 2017. The high, middle and low tiers of residential homes' year over year growth rates by December 2017 are 4.25%, 7.28%, and 5.9%. Using the seasonal ARIMA model, the predicted year over year growth rates for 3 tiers by the end of 2018 are 3.54% for the high tier, 6.73% for the middle tier and 6.07% for the low tier such that the middle tier is the main growth factor for New York market. It can be seen that the growth rate for both the high and middle tiers in 2018 will be inferior to that in 2017, but the low tier market is gonna see a slightly steeper growth in 2018.

```{r}
knitr::opts_chunk$set(warning = FALSE)

# convert San Francisco Home Index data (3 tiers) into time series object
sfxrhtnsats <- ts(SFXRHTNSA[,2], start = c(1987, 1), frequency = 12)
sfxrltnsats <- ts(SFXRLTNSA[,2], start = c(1987, 1), frequency = 12)
sfxrmtnsats <- ts(SFXRMTNSA[,2], start = c(1987, 1), frequency = 12)

# trim the time series objects to be a multiplier of 12 monthes
sfxrhtnsats <- yearTrim(sfxrhtnsats)
sfxrltnsats <- yearTrim(sfxrltnsats)
sfxrmtnsats <- yearTrim(sfxrmtnsats)

# predict for high tier
pred.sfht <- nextYearPred(sfxrhtnsats)$mean

# predict for low tier
pred.sflt <- nextYearPred(sfxrltnsats)$mean

# predict for middle tier
pred.sfmt <- nextYearPred(sfxrmtnsats)$mean

# combine the original data and the prediction into one data frame 
sfxrnsa <- cbind(sfxrhtnsats, sfxrmtnsats, sfxrltnsats, pred.sfht, pred.sfmt, pred.sflt)

# use dygraphs package to plot the data and the prediction for three tiers in one graph
dygraph(sfxrnsa, main = "Home Price Index for San Francisco, California") %>%
  dyAxis("x", label = "Time", drawGrid = FALSE, pixelsPerLabel = 40, axisLabelFormatter = "function(d){return d.getFullYear();}") %>%
  dyAxis("y", label = "Price Index") %>%
  dySeries("sfxrhtnsats", label = "High Tier", strokeWidth = 3) %>%
  dySeries("sfxrmtnsats", label = "Middle Tier", strokeWidth = 3) %>%
  dySeries("sfxrltnsats", label = "Low Tier", strokeWidth = 3) %>%
  dySeries("pred.sfht", label = "Prediction for High Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dySeries("pred.sfmt", label = "Prediction for Middle Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dySeries("pred.sflt", label = "Prediction for Low Tier", strokeWidth = 4, strokePattern = c(7, 3)) %>%
  dyShading(from = "2018-1-1", to = "2018-12-1") %>%
  dyOptions(strokeWidth = 3, colors = RColorBrewer::brewer.pal(3, "Dark2")) %>%
  dyLegend(width = 600)

# compare the index for all three tiers between the most current and one years ago
compareInd(sfxrhtnsats, c(2016,12), 12)
compareInd(sfxrmtnsats, c(2016,12), 12)
compareInd(sfxrltnsats, c(2016,12), 12)

# compare the predicted index in 2018.12 and the index in 2017.12
(pred.sfht[12] - sfxrhtnsats[372])/ sfxrhtnsats[372]
(pred.sfmt[12] - sfxrmtnsats[372])/ sfxrmtnsats[372]
(pred.sflt[12] - sfxrltnsats[372])/ sfxrltnsats[372]
```
It can be observed that the home price indexes for all 3 tiers in San Francisco, CA have experienced significant growth. Although the price index for low tier hasn't reached the highest mark before 2008 financial crisis by the end of 2017, the rest tiers have already exceeded. The high, middle and low tiers of residential homes' year over year growth rates by December 2017 are 8.38%, 10.06% and 10.98%, which are very high index across the nation. Using the seasonal ARIMA model, the predicted year over year growth rates for 3 tiers by the end of 2018 are 6.96% for the high tier, 15.29% for the middle tier and 13.68% for the low tier. It can be seen that the growth rate for both the high and middle tiers in 2018 will be much higher to that in 2017, and both the middle and low tier is gonna experience a bigger blooming in 2018. The future of San Francisco market seems to be very bright.